---
title: "Example 04: Tables and dplyr"
engine: knitr
---

```{r}
#| echo: false

knitr::opts_chunk$set(engine.opts = list(zsh = "-l"))
```



# Crosstabulations

```{r}
library(tidyverse)
library(socviz)

gss_sm
```


Count up one variable:

```{r}
gss_sm |>
  group_by(bigregion) |>
  tally()
```


Cross-tabulate:

```{r}
gss_sm |>
  group_by(bigregion, religion) |>
  tally()
```

## Groups are nested; order matters

Notice the difference:

```{r}
gss_sm |>
  group_by(religion, bigregion) |>
  tally()
```

The Ns are the same but the logical grouping is not. Now try:

```{r}
# Religion within bigregion;
gss_sm |>
  group_by(bigregion, religion) |>
  tally() |>
  summarize(group_total = sum(n)) |>
  mutate(prop = group_total/sum(group_total))
```


```{r}
# Bigregion within religion
gss_sm |>
  group_by(religion, bigregion) |>
  tally() |>
  summarize(group_total = sum(n)) |>
  mutate(prop = group_total/sum(group_total))
```


To get our table to look like a conventional n by m crosstab, pivot it:

```{r}
gss_sm |>
  group_by(bigregion, religion) |>
  tally() |>
  pivot_wider(names_from = religion, values_from = n)
```

We'll learn more about pivoting next week.

NB There are functions and whole packages dedicated to making crosstabs (including tidy crosstabs) just with a single
simplified call, e.g.
[`janitor::tabyl()`](https://cran.r-project.org/web/packages/janitor/vignettes/tabyls.html)
or more extravagantly [`crosstable`](https://danchaltiel.github.io/crosstable/).
But that's not our focus here.


## Behavior of groups with `summarize()` and `mutate()`

```{r}
# Bigregion within religion
gss_sm |>
  group_by(religion, bigregion) |>
  tally()
```


Consider the difference between this:

```{r}
# Bigregion within religion
gss_sm |>
  group_by(religion, bigregion) |>
  tally() |>
  # Summarize
  summarize(group_total = sum(n))
```


And this:

```{r}
# Bigregion within religion
gss_sm |>
  group_by(religion, bigregion) |>
  tally() |>
  # Mutate
  mutate(group_total = sum(n))
```


As many nested grouping dimensions as we wish:

```{r}
gss_sm |>
  group_by(bigregion, race, sex) |>
  tally()
```


Equivalently, but the result is *not* grouped:

```{r}
gss_sm |>
  count(bigregion, race, sex)
```


Add a frequency column:

```{r}
gss_sm |>
  group_by(bigregion, race, sex) |>
  tally() |>
  mutate(prop = n/sum(n))
```


When the result is not grouped, what do you get as the proportions?

```{r}
gss_sm |>
  count(bigregion, race, sex) |>
  mutate(prop = n/sum(n))
```

Check your work by summing the rows.
