---
title: "Example 05: More dplyr and tidyr"
engine: knitr
---

```{r}
#| echo: false

knitr::opts_chunk$set(engine.opts = list(zsh = "-l"))
```


```{r}
library(tidyverse)
library(nycdogs)
library(socviz)
```

## Using `across()` to apply functions to multiple columns

```{r}
gss_sm
```

Let's summarize the age, childs, and sibs columns.

```{r}
gss_sm |>
  summarize(
    age_mean = mean(age, na.rm = TRUE),
    age_sd = sd(age, na.rm = TRUE),
    childs_mean = mean(childs, na.rm = TRUE),
    childs_sd = sd(childs, na.rm = TRUE),
    sibs_mean = mean(sibs, na.rm = TRUE),
    sibs_sd = sd(sibs, na.rm = TRUE)
  )
```

This is tedious and we have to name all the columns we create, which is error-prone. We can use `across()` to apply the same functions to multiple columns at once.

```{r}
gss_sm |>
  summarize(
    across(c(age, childs, sibs),
           list(mean = \(x) mean(x, na.rm = TRUE),
                sd = \(x) sd(x, na.rm = TRUE)))
  )
```

We can also use `where()` to select columns based on their type. Here we summarize all numeric columns. First we drop `id` because it's just a row-identifier and it doesn't make sense to summarize it. We drop `year`, `ballot`, and the weight variables for similar reasons. We use a tidy selector to find the weight variables, and negate with `-` before the `c()` to drop the slected columns. Then in the summarize statement we use `where(is.numeric)` to select all remaining columns that are numeric.

```{r}
gss_sm |>
  select(-c(id, ballot, year, starts_with("wts"))) |>
  summarize(
    across(where(is.numeric),
           list(mean = \(x) mean(x, na.rm = TRUE),
                sd = \(x) sd(x, na.rm = TRUE)))
  )
```

This will of course work with grouping variables as well. Here we also add a count of the number of rows in each group:

```{r}
gss_sm |>
  select(-c(id, ballot, year, starts_with("wts"))) |>
  group_by(bigregion, degree) |>
  summarize(
    n = n(),
    across(
      where(is.numeric),
      list(mean = \(x) mean(x, na.rm = TRUE), sd = \(x) sd(x, na.rm = TRUE))
    )
  ) |>
  print(n = Inf) # n = Inf means "Print every row to the console"
```

See how, when we do this, we don't put the `n()` call in the `across()`? That's
because `n()` is not a function that we are applying across all the numeric
variables. We're just counting the number of rows within the group, which will
not change no matter how many summary statistics we calculate for variables
within each group.

## Per-Operation Grouping

As an aside, consider the difference between doing it with a separate
`group_by()` line, as above, and doing the same thing with `summarize(.by =
c(...) ...)`:

```{r}
gss_sm |>
  select(-c(id, ballot, year, starts_with("wts"))) |>
  summarize(
    .by = c(bigregion, degree),
    n = n(),
    across(
      where(is.numeric),
      list(mean = \(x) mean(x, na.rm = TRUE), sd = \(x) sd(x, na.rm = TRUE))
    )
  ) |>
  print(n = Inf)
```


The values for any particular row are the same, but the row-order may not be.


## Summarizing columns while pivoting

```{r }
#| label: "04-tidy-data-19"

# Some made-up data
dfstrat <- read_csv(here::here("data", "dfstrat.csv"))
dfstrat
```

This is a simulated sample of a 1,000 people in eight survey strata, with
measured sex, race, and education characteristics. For simplicity we just have
two categories within each measure. We have an income measure for each person.
So we have e.g. sex x race x educ combinations of categories and we want to know
things about average income within various categories.

Let’s say we want to transform this to a wider format, specifically by widening
the `educ` column, so we end up with columns for both the HS and BA categories.
While we do so we also want to calculate both the mean of income and the total n
within each group, still within strata, while holding on to the sex and race
classifcations. We "over-group" the sex and race categories, and then have
education within stratum:

```{r}
dfstrat |>
  group_by(sex, race, stratum, educ) |>
  summarize(
    mean_inc = mean(income),
    n = n()
  ) |>
  pivot_wider(names_from = (educ), values_from = c(mean_inc, n)) |>
  ungroup()
```

Here we end up with sex-by-race-by-stratum in the rows, and the income-by-education means, and income-by-education Ns, in their own columns.

Again, consider the difference between that and the `.by ...` way:

```{r}
#| label: "04-tidy-data-20"
dfstrat |>
    summarize(.by = c(sex, race, stratum, educ),
              mean_inc = mean(income),
              n = n()) |>
    pivot_wider(names_from = (educ),
                values_from = c(mean_inc, n))
```

The values in any particular cells are the same, but the row order may be
different, and also the column order (thanks to the way the pivot interacts with
the per-operation grouping).
