---
title: "Example 07: Functions and Iteration"
engine: knitr
---

```{r}
#| echo: false

knitr::opts_chunk$set(engine.opts = list(zsh = "-l"))
library(tidyverse)
library(here)
library(socviz)

## ---
## This code is not necessary to reproduce the example
library(myriad)
import_myriad_semi()
import_myriad_condensed()
theme_set(theme_myriad_semi())
## ---

```

```{r}
library(broom)
```


## Writing Functions

I mentioned in class that my add function didn't have any error handling or
guard clauses. Here's an example of why it's worth adding them.

```{r}
my_adder <- function(x, add = 1) {
  x + add
}

my_adder(10)

my_adder(1:10)

# This won't work
# my_adder("a")
```

Guard clauses and error handling:

```{r}
my_adder <- function(x, add = 1) {
  if(!is.numeric(x)) {
    return("Argument `x` must be numeric.")
  }

  x + add
}

my_adder("a")


my_adder <- function(x, add = 1) {
  if(!is.numeric(x)) {
    return("Argument `x` must be numeric.")
  }

  if(!is.numeric(add)) {
    return("Argument `add` must be numeric.")
  }

  x + add
}


```

Here I use `return()` in the guard clauses in order to make this example work as
part of the pipeline that creates the site. But really I should use `stop()`.

```{r}
my_adder(1:10, 5)


my_adder(1:10, "c")
```


One small point here: resist using `else` in clauses like this. Just use `if` to
check a condition and if it fails, exit the function. The "happy path" is one
that passes your guard clauses. Other paths should exit with an informative
message.
