---
title: "README File"
format: gfm
---

```{r}
#| include: false

## Using the {{< var >}} shortcode doesn't work in the README for some reason. 
yaml_vars <- yaml::read_yaml(here::here("_variables.yml"))
```


<!-- README.md is generated from README.qmd. Please edit that file -->


# `r yaml_vars$course$number`: `r yaml_vars$course$title` 

- Taught by `r yaml_vars$instructor$name`, `r yaml_vars$university$name`
- `r yaml_vars$course$semester`

This is the source code for the course website at <`r yaml_vars$course$url`>

The site is a [Quarto](https://quarto.org) project. It uses [{{renv}}](https://rstudio.github.io/renv/) to manage dependencies and [{{targets}}](https://docs.ropensci.org/targets/) to manage the site building. 


To build the site from scratch, first you have to ~create the universe~ open the project in RStudio by clicking on `mptc.Rproj`. 

- If it’s not installed already, R should try to install the {renv} package when you open the RStudio Project for the first time. 
- If you don’t see a message about package installation, install it yourself by running install.packages("renv") at the R console.
- Run renv::restore() in the R console to install all the required packages for this project.
- Run targets::tar_make() in the R console to build everything.


## Everything is a DAG now

Here is the dependency graph for the site build.

```{r}
#| echo: false
#| eval: true
#| collapse: true

withr::with_dir(here::here(), {
  targets::tar_load(workflow_graph)
})
```

```{r}
#| echo: false
#| results: asis

# Make sure there are newlines between the code block boundaries and the output
# from `workflow`, which also needs to be separated by newlines
cat("``` mermaid\n")
cat(workflow_graph, sep = "\n")
cat("\n```")
```

