---
title: "01. Doing Your Work Properly"
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

Why is technical computing, in the broadest sense, organized in the way that it is? The specifics are, inevitably, the path-dependent product of particular historical developments. But we can also motivate the topic more positively in terms of ideas about organization and control through composable actions on files. We need to repeatedly do things with the computer; we want to be able to say and have some record of what we did; and we want to be able to open up and inspect the actions we took in order to change them if needed.

## Reading

PDFs of all these readings are available on Canvas.

- @christensen19:_trans_reprod_social_scien_resear "Introduction" and Chapter 11, "Reproducible Workflow".
- @healy20:_plain_person_guide_plain_text_social_scien "1. Introduction"
- @lewis-kraus2023dishonesty
- @lewis-kraus2024defamation
- @crimson2025gino

## Other work

-   Familiarize yourself with the [content](/content/index.qmd), [examples](/example/index.qmd), and [assignments](/assignment/index.qmd) pages for this class.


## Assignments

- This week (and only this week) there are _two_ assignments. [Assignment 00](/assignment/00-assignment-install-r-github.qmd) is, as it were, the zeroth assignment and is due on the first day of class. [Assignment 01](/assignment/01-assignment-quarto-notes.qmd) is the first regular assignment and is due Friday, August 29th.

## Slides


```{=html}
<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>
<iframe class="slide-deck" id="quartoframe" src="../slides/01-slides-bigpicture.html"></iframe


```
```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/01-slides-bigpicture")
```


```{=html}
<iframe class="slide-deck" id="quartoframe" src="../slides/01-slides-rstudio-firstlook.html"></iframe
```

```{r show-slide-tabs2, echo=FALSE, results="asis"}
slide_buttons("/slides/01-slides-rstudio-firstlook")
```
