---
title: "03. Using R to look at data"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

We'll spend most of the rest of the semester working in R. Time to learn it! We
will learn some basics of R mostly via drawing graphs of our data.

## Required Reading

- @wickham23:_r_data_scien.

[Read the text online.](https://r4ds.hadley.nz/) Complete as much of the "[Whole
Game](https://r4ds.hadley.nz/whole-game)" section as you can, i.e. Chapters 1-8.
(They are short.) Use the R project for your chosen data, create a new project,
or add a Quarto document to your existing class notes project, and try doing
some of the examples and code in the Whole Game section using data of your own,
or the dataset you chose from last week's assignment. Don't worry if things
don't go as expected, but do make a note of what you found difficult or
confusing so we can discuss it in seminar.

- @healyDataVisualizationPractical2027 Read Chapters 1 to 4. There's a good deal
of overlap with Wickham et al and with the slides.

I've also put the slides up in advance, as we will be working more interactively
in class starting this week. So work through the slides alongside the reading.
Again, there's a subtantial amount of overlap between the readings and the
slides. Hopefully the repetition will help you get the gist of how the language
works. Practice is the best way to learn, though. Ask questions in the Slack if
you run into anything you don't understand, or come to class ready to ask.

## Examples

[R basics and ggplot examples.](/example/03-example-ggplot.qmd)

## Assignment


## Slides

The "R and Quarto" slides repeat some of the material from the first week, just
for convenience, and then move on to the introduction to R and ggplot.

```{=html}
<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>
<iframe class="slide-deck" id="quartoframe" src="../slides/03-slides-r-and-quarto.html"></iframe>
```

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/03-slides-r-and-quarto")
```


```{=html}
<iframe class="slide-deck" id="quartoframe" src="../slides/03-slides-ggplot-1.html"></iframe>
```

```{r show-slide-tabs2, echo=FALSE, results="asis"}
slide_buttons("/slides/03-slides-ggplot-1")
```


```{=html}
<iframe class="slide-deck" id="quartoframe" src="../slides/03-slides-ggplot-2.html"></iframe>
```

```{r show-slide-tabs3, echo=FALSE, results="asis"}
slide_buttons("/slides/03-slides-ggplot-2")
```
