---
title: "04. Using R to look at data"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

We'll spend most of the rest of the semester working in R and RStudio. Time to learn it!

## Required Reading

- @wickham23:_r_data_scien.

[Read the text online.](https://r4ds.hadley.nz/) Complete as much of the "[Whole Game](https://r4ds.hadley.nz/whole-game)" section as you can, i.e. Chapters 1-8. (They are short.) Create a new R project, or add a Quarto document to your existing class notes project, and try redoing some of the examples and code in the Whole Game section using data of your own, or the dataset you chose from last week's assignment. Don't worry if things don't go as expected, but do make a note of what you found difficult or confusing so we can discuss it in seminar.



## Examples

[R basics and ggplot examples](/example/04-example-ggplot.qmd)

## Assignment

No assignment this week.


## Slides

The "R and Quarto" slides repeat some of the material from the first week, just for convenience, and then move on to the introduction to R. There is more material in both slide decks than we can covered in class.

```{=html}
<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>
<iframe class="slide-deck" id="quartoframe" src="../slides/04-slides-r-and-quarto.html"></iframe>
```

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_win_button("/slides/04-slides-r-and-quarto")
```


```{=html}
<iframe class="slide-deck" id="quartoframe" src="../slides/04-slides-ggplot-1.html"></iframe>
```

```{r show-slide-tabs2, echo=FALSE, results="asis"}
slide_win_button("/slides/04-slides-ggplot-1")
```
