---
title: "4. R and Quarto"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

We'll spend most of the rest of the semester working in R and RStudio. Time to learn it!

## Required Reading

- @wickham23:_r_data_scien. 

[Read the text online.](https://r4ds.hadley.nz/) Complete as much of the "[Whole Game](https://r4ds.hadley.nz/whole-game)" section as you can, i.e. Chapters 1-8. (They are short.) If you can, see if you can create a new R project and try redoing some of the examples and code in the Whole Game section using data of your own, or a dataset downloaded from one of the [practice data sources](https://mptc.io/content/). Don't worry if things don't go as expected, but do make a note of what you found difficult or confusing so we can discuss it in seminar.



## Examples

[This week's examples and demos](/example/04-example.qmd).


## Assignment

[This week's assignment](/assignment/04-assignment.qmd) is essentially the reading, done as actively as you can.

## Slides

```{=html}
<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>
<!--iframe class="slide-deck" id="quartoframe" src="../slides/03-slides.html"></iframe-->
```

```{r show-slide-tabs, echo=FALSE, results="asis"}
#slide_win_button("/slides/03-slides")
```
