---
title: "05. Tidying, Reshaping, and Ingesting Data"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

More on dplyr: using  `across()` to apply functions to multiple columns at once;
using `case_when()` to create new variables based on conditions; reshaping data
with `pivot_longer()`;  importing and cleaning your data from plain-text and
foreign binary formats.

## Required Reading

- @bromanDataOrganizationSpreadsheets2018
- @wickham23:_r_data_scien. Read [Chapter 7](https://r4ds.hadley.nz/data-import) on importing data. Read [Chapter 20](https://r4ds.hadley.nz/spreadsheets) on Spreadsheets. Read or review [Chapter 17](https://r4ds.hadley.nz/datetimes.html) on dates and times.

### `tidyr`

- [Tidy data](https://tidyr.tidyverse.org/articles/tidy-data.html)
- [Pivoting](https://tidyr.tidyverse.org/articles/pivot.html)
- (Optional.) [Rectangling](https://tidyr.tidyverse.org/articles/rectangle.html)

### `stringr`

- [Regular expressions](https://stringr.tidyverse.org/articles/regular-expressions.html)
- Read the help page for `str_detect()`: `?stringr::str_detect`, which can also be found [here](https://stringr.tidyverse.org/reference/str_replace.html). Work through the examples one at a time and make a note of any that seem confusing.


## Examples

[This week's examples.](/example/05-example-tidying-and-ingesting.qmd)


## Assignment

[This week's assignment.](/assignment/05-assignment-fars.qmd)

## Slides

```{=html}
<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>
```

### Tidying and reshaping data

```{=html}
<iframe class="slide-deck" id="quartoframe" src="../slides/05-slides-r-reshaping-data.html"></iframe>
```

```{r show-slide-tabs2, echo=FALSE, results="asis"}
slide_buttons("/slides/05-slides-r-reshaping-data")
```

### Regular Expressions intro

```{=html}
<iframe class="slide-deck" id="quartoframe" src="../slides/05-slides-r-search-and-edit-text.html"></iframe>
```

```{r show-slide-tabs3, echo=FALSE, results="asis"}
slide_buttons("/slides/05-slides-r-search-and-edit-text")
```

### Ingesting data

```{=html}
<iframe class="slide-deck" id="quartoframe" src="../slides/05-slides-r-ingesting-data.html"></iframe>
```

```{r show-slide-tabs4, echo=FALSE, results="asis"}
slide_buttons("/slides/05-slides-r-ingesting-data")
```
