---
title: "6. Version Control and Git"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

We'll learn about backing up your work and keeping the most important parts in a version control sytem.

## Required Reading

-   @bryanHappyGit Read *and work through* as much of this book as you can, but at least Chapters 1, 4 through 9, 11, and 12.

## Optional but Recommended

-   @evansLearningGit2024

-   @sylor-miller21:_oh_shit_git.

## Optional

-   @chacon14:_pro_git

## Examples

[This week's examples and demos](/example/06-example.qmd).

## Assignment

[This week's assignment](/assignment/06-assignment.qmd) is to follow the tasks outlined in *Happy Git With R*, to the point where you have a sample repo you have committed to and pushed.

## Slides

<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>

<iframe class="slide-deck" id="quartoframe" src="../slides/06-slides.html"></iframe>

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_win_button("/slides/06-slides")
```
