---
title: "06. More on dplyr and reshaping data"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

More on dplyr: using  `across()` to apply functions to multiple columns at once; using `case_when()` to create new variables based on conditions; more on reshaping data with `pivot_longer()`; more on filtering and selecting data.

## Required Reading

- @bromanDataOrganizationSpreadsheets2018

### `tidyr`

- [Tidy data](https://tidyr.tidyverse.org/articles/tidy-data.html)
- [Pivoting](https://tidyr.tidyverse.org/articles/pivot.html)
- (Optional.) [Rectangling](https://tidyr.tidyverse.org/articles/rectangle.html)

### `stringr`

- [Regular expressions](https://stringr.tidyverse.org/articles/regular-expressions.html)
- Read the help page for `str_detect()`: `?stringr::str_detect`, which can also be found [here](https://stringr.tidyverse.org/reference/str_replace.html). Work through the examples one at a time and make a note of any that seem confusing.


## Examples

[This week's examples.](/example/06-example-dplyr-and-more.qmd)


## Assignment

## Slides

```{=html}
<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>
```

```{=html}
<iframe class="slide-deck" id="quartoframe" src="../slides/05-slides-r-reshaping-data.html"></iframe>
```

```{r show-slide-tabs2, echo=FALSE, results="asis"}
slide_buttons("/slides/05-slides-r-reshaping-data")
```
