---
title: "7. Getting data in and out of R"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

We'll learn about importing and cleaning your data.

## Required Reading

- @wickham23:_r_data_scien. 

[Read the text online.](https://r4ds.hadley.nz/). Read or review [Chapter 7](https://r4ds.hadley.nz/data-import) on importing data. Read [Chapter 20](https://r4ds.hadley.nz/spreadsheets) on Spreadsheets. Read and work through [Chapters 12 to 18](https://r4ds.hadley.nz/transform) in the "Transform" part of the book.


## Examples

[This week's examples and demos](/example/07-example.qmd).

## Assignment

[This week's assignment](/assignment/07-assignment.qmd).

## Slides

<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>

<iframe class="slide-deck" id="quartoframe" src="../slides/07-slides.html"></iframe>

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_win_button("/slides/07-slides")
```
