---
title: "8. Databases"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

Working with data that's nominally larger than the amount of RAM you have.

## Required Reading

- @wickham23:_r_data_scien. 

[Read the text online.](https://r4ds.hadley.nz/). Take a look at [Chapter 21](https://r4ds.hadley.nz/databases) on databases. data. It uses a different database back-end from the one we'll use in class, but the ideas are the same.

## Examples

[This week's examples and demos](/example/08-example.qmd).

## Assignment

No assignment this week.

## Slides

<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>

<iframe class="slide-deck" id="quartoframe" src="../slides/08a-slides.html"></iframe>

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_win_button("/slides/08a-slides")
```
