---
title: "09. Iterating on Data and Models"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

Repeatedly doing things to data; repeatedly doing things to models. Core functional programming patterns.

## Reading

- @wickham23:_r_data_scien.

[Read the text online](https://r4ds.hadley.nz/). Read [Chapter 25](https://r4ds.hadley.nz/functions) on functions and [Chapter 26](https://r4ds.hadley.nz/iteration) on iteration.


## Examples

[This week's examples and demos](/example/09-example-iteration.qmd).

## Assignment

## Slides

<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>

<!--iframe class="slide-deck" id="quartoframe" src="../slides/08-slides-git.html"></iframe-->

```{r show-slide-tabs, echo=FALSE, results="asis"}
#slide_buttons("/slides/08-slides-git")
```
