---
title: "10. Supercharged Iteration: Parallel Computing"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

Sometimes, embarrassing is good. Like when a task is embarrassingly parallel.

## Required Reading

- The Futureverse: [https://www.futureverse.org](https://www.futureverse.org)
- The `furrr` package website: [https://furrr.futureverse.org](https://furrr.futureverse.org)
- Henrik Bengtsson's "[Futureverse - A Unifying Parallelization Framework in R for Everyone](https://henrikbengtsson.github.io/course-stanford-futureverse-2023/)" workshop



## Examples

[This week's examples and demos](/example/10-example.qmd).

## Assignment

(No assignment this week.)

## Slides

<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>

<iframe class="slide-deck" id="quartoframe" src="../slides/10-slides.html"></iframe>

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_win_button("/slides/10-slides")
```
