---
title: "12. Build Systems, Environments, Packages"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

We will look at tools that help organize your work into projects, manage dependencies, and share your code with others.


## Assignment

There's an assignment due next Monday. [Here it is.](/assignment/12-assignment-enchilada.qmd)

## Slides

<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>

<!--iframe class="slide-deck" id="quartoframe" src="../slides/11-slides-build.html"></iframe-->

```{r show-slide-tabs, echo=FALSE, results="asis"}
#slide_win_button("/slides/11-slides-build")
```
