---
title: "13. Packages and LLMs"
date: last-modified
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

Getting into the habit of packaging your work. Also, what LLMs can and cannot do for you right now.

## Reading

### Packages
- @wickhambryan:2023, Introduction; Chapters 1-11, 16.

### AI and LLMs
- Vicki Boykis (2025). [How big are our embeddings now and why?](https://vickiboykis.com/2025/09/01/how-big-are-our-embeddings-now-and-why/)
- Glyph (2025). [I Think I’m Done Thinking About genAI For
Now.](https://blog.glyph.im/2025/06/i-think-im-done-thinking-about-genai-for-now.html)
_Deciphering Glyph_.
- Rob Napier (2025). [TIL:AI. Thoughts on
AI](https://cocoaphony.micro.blog/2025/06/16/tilai-thoughts-on-ai.html). *Cocoaphony*.
- Thomas Ptacek (2025). [My AI Skeptic Friends Are All
Nuts](https://fly.io/blog/youre-all-nuts/). Fly.io.
- Tristan Mahr (2025). [readtextgrid now uses C++ (and ChatGPT helped)](https://www.tjmahr.com/readtextgrid-cpp-llms/).
- Ellmer Vignettes:
    - [Programming with
    Ellmer](https://ellmer.tidyverse.org/articles/programming.html)
    - [Prompt Design](https://ellmer.tidyverse.org/articles/prompt-design.html)
    - [Structured Data
    Basics](https://ellmer.tidyverse.org/articles/structured-data.html)
    - [Tool/Function Calling](https://ellmer.tidyverse.org/articles/tool-calling.html)

## Slides

<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>

<!--iframe class="slide-deck" id="quartoframe" src="../slides/12-slides.html"></iframe-->

```{r show-slide-tabs2, echo=FALSE, results="asis"}
#slide_win_button("/slides/12-slides")
```
