---
title: "Week 05 Problem Set: Dogs of New York"
author: "<PUT YOUR NAME HERE>"
date: "`r Sys.Date()`"
format: html
---

```{r}
#| label = "load-libraries-data",
#| warning = FALSE,
#| message = FALSE
library(tidyverse)
```

# Dogs of New York

This data comes from New York City's Open Data initiative.

We load the data as a package. If you get a message saying there's no such package, then run this line of code at the console _once_:  `remotes::install_github("kjhealy/nycdogs")`. Then it should be available.

```{r}
library(nycdogs)

nyc_license
```

The data are structured like this: each _row_ contains the license record for a dog in a particular year. The columns are as follows:

`animal_name` --- Name of dog.
`animal_gender` --- Sex of dog.
`animal_birth_year` --- Birth year
`breed_rc` --- Breed of dog (simplier categories).
`borough` --- Borough of owner. (Some boroughs are approximate due to zip codes.)
`zip_code` --- Zip code of owner.
`license_issued_date` --- Date license issued
`license_expired_date` --- Date license expired
`extract_year` --- Year record was extracted


# Tasks

## 1. A first look at the data.

You'll use `group_by()` and `summarize()`, or `tally()` to answer these questions.

- a. Write some code to count the number of records there are for each extract year, and now many records are missing a value for the extract year.

```{r}

```

- b. How many distinct dog breeds are there in the data?


- c. Are there any notable differences in the ratio of male to female dogs in the city as a whole? What about by borough?


- d. Here are the first ten rows of the data again:

```{r}
nyc_license
```

Row 8 shows a Labrador named Chewbacca. Are there any other dogs in the data that have the names of major Star Wars characters?

- e. In class we saw some problems with the `animal_name` column, where there seemed to be typos and misspellings and so on. Are there any similar sorts of problems with the `borough` column or the `zip_code` column? (Hint: remember we subset rows with `filter()`)

## 2. Some Crosstabs

- a. What were the Top 5 dog names in each borough in 2016?

```{r}

```

- b. What were the Top 5 dog breeds in the whole city in 2018, in terms of the total number of dogs?

- c. What were the Top 5 dog breeds in the whole city in 2018, as a percentage of all dogs?

- d. What were the Top 5 dog breeds in each borough in 2018, in terms of the total number of dogs in each borough?

- c. Redo this table to show the top 5 dog breeds in each borough in 2018, this time as a percentage of all dogs in the borough.

## 3. A Graph

- a. Make a table of the top 10 dog *names* in each borough in 2018

- b. Make a graph with this table. First think about how best to graph it. Consider either `geom_col()` and `geom_point()`. Add a title and label the axes. Facet the graph by borough.

- c. Can you make it so that only the dog names in each borough are shown for that borough's facet?
