---
title: "3. Text Editing"
date: 2023-09-12
date-format: full
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## This week ...

Most of our work is not done interactively. We write things down as "code" in its broadest sense: just sequences of instructions and commentary on them. We need a good set of tools for editing and manipulating all the text we produce. So this week we will cover

- A little more on shell scripting
- The idea of text search and manipulation via regular expressions
- Text Editors

## Background and Related Reading

- @robbins05:_class_shell_scrip
- @neil15:_pract_vim
- @petersen22:_master_emacs
- @friedl06:_master_regul_expres

## Examples

[This week's examples and demos](/example/03-example.qmd).


## Slides

```{=html}
<script type="application/javascript">

function resizeIFrameToFitContent( iFrame ) {

   iFrame.width  = document.getElementById("quarto-document-content").clientWidth;
   iFrame.height  = (document.getElementById("quarto-document-content").clientWidth * 0.5625)
}

window.addEventListener('DOMContentLoaded', function(e) {

    var iFrame = document.getElementById( 'quartoframe' );
    resizeIFrameToFitContent( iFrame );
    } );

</script>
<iframe class="slide-deck" id="quartoframe" src="../slides/03-slides.html"></iframe>
```

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_win_button("/slides/03-slides")
```
