---
title: "Sociol 703: Modern Plain Text Computing" 
author:
  - name: Kieran Healy
    email: kieran.healy@duke.edu
    affiliations: 
      - id: duke
        name: Duke University
format:
  html: default
  minion-pdf: default

---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tibble)
library(pander)
```

# About this course

This course is an introduction to modern plain-text methods of data analysis, data management, and coding. It is requried for first-year graduate students in the Sociology department. It introduces a suite or "stack" of computing tools and techniques that are widely used in the department, the discipline, across the social sciences, and beyond. We will learn to use these tools and also learn why they exist and why they are important for producing work that is reliable, reproducible, and open to inspection. 

# Motivation 

Researchers depend on computer software to get their work done. But often, they do not know enough about how their computers work. This makes their lives more difficult. I don't mean it's a shame not everyone is a software engineer ready and able to write applications from the ground up. Rather, as researchers working with data of various kinds, we just don't make the best use of our computers. Nor are we encouraged to reflect on why they work the way they do. Students end up fending for ourselves and, hopefully, picking things up informally. Instructors are forced to spend time bringing people up to speed about where that document went, or what a file is, or why the stupid thing stopped working just now. In the worst case, we never properly understand these tools at all and end up marinating in an mixture of magical thinking and sour resentment towards the machines we sit in front of for hours each day.

The fragmented, messy, and multifacted tasks associated with scholarly research make heavy demands on software. Most of them have to do with the need for *control* over what you are doing, and especially the importance of having a *record* of what you did that you can revisit and *reproduce* if needed. Our tools also need to allow us to track down and diagnose errors. Because research in practice is fragmented and messy, this can often be a tricky process to think clearly about and work through in a systematic way.

To help address these challenges, modern computing platforms provide us with a suite of powerful, modular, specialized tools and methods. The bad news is that they are not magic; they cannot do our thinking for us. The good news is that they are by now well-established and stable. Most of them are developed in the open. Many are supported by helpful communities. Almost all are available for free. Nearly without exception, they tend to work through the medium of explicit instructions written out in plain text. In other words they work by having you write some code, in the broadest sense. People who do research involving structured data of any kind should become familiar with these tools. Lack of familiarly with the basics encourages bad habits and unhealthy attitudes among the informed and uninformed alike.



# Reading

Required reading from books and articles will be provided on [the course website](https://mptc.io) Canvas or (in most cases) will be freely available online. Useful texts to acquire in hardcopy or to bookmark include:

- [@wickham23:_r_data_scien] 
- [@janssens21:_data_science_command_line]
- [@ismaykim19:modern]
- [@chacon14:_pro_git]
- [@healy:2019] 
- [@friedl06:_master_regul_expres]
- [@landau22:_r]
- [@wickhambryan:2023]
- [@christensen19:_trans_reprod_social_scien_resear]

# Software

We will do all most of our work class using Unix-style command line tools, 
[R](https://cran.r-project.org/), and  [RStudio](https://www.rstudio.com/). [R](https://cran.r-project.org/) is a freely-available programming language that is designed for statistical computing and widely used across the natural and social sciences, as well as in the world of "data science" generally. [RStudio](https://www.rstudio.com/) is an integrated development environment, or IDE, for R, a kind of control center from which you can manage the engine-room
of R itself. It is also freely available. 

# Course website

The course website is at <https://mptc.io>. Each week, slides, readings, and other class material will be posted there along with additional examples and the problem set due the following week.

# Weekly Schedule

```{r,echo=FALSE}
readxl::read_xlsx(here::here("data", "schedule.xlsx")) |> 
  dplyr::mutate(date = lubridate::ymd(date), 
                day2 = lubridate::ymd(day2),
                date = glue::glue('{format(date, "%b %e")} / {format(day2, "%b %e")}'), 
                date = stringr::str_replace_all(date, "NA", "-")) |> 
  dplyr::select(subgroup, title) |> 
  dplyr::rename(Week = subgroup, Topic = title) |> 
  kableExtra::kbl(booktabs = TRUE)
```


# Course policies

- Attendance is required, and important. I am a reasonable person; if you need to be absent please _let me know in advance_ insofar as that is possible.
- Do the assigned readings in advance of class.
- Submit problem sets, or other assignments, on time.

# Required work and grading

Weekly **Class Participation** (30% of final grade) and **Problem Sets** (50% of final grade) will let you reflect on the reading and practice your skills. Problem sets are due by end of day the _Monday_ after they are assigned. A **Final Project** (20% of final grade) will allow you to put what you've learned to use. 


# Duke community standard

Like all classes at the university, this course is conducted under the Duke
Community Standard. Duke University is a community dedicated to scholarship,
leadership, and service and to the principles of honesty, fairness, respect, and accountability. Citizens of this community commit to reflect upon and uphold these principles in all academic and nonacademic endeavors, and to protect and promote a culture of integrity. To uphold the Duke Community Standard you will
not lie, cheat, or steal in academic endeavors; you will conduct yourself
honorably in all your endeavors; and you will act if the Standard is
compromised.
